---
title: "Assignment_3"
author: "Sara Szabo"
date: "2026-04-28"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Task description

In assignment 3 you have to analyze real world data using Bayesian models of cognition. You can apply the models we discussed during the lectures, but you need to adjust them to reflect the specific setup.

The data comes from the social conformity experiment (<https://pubmed.ncbi.nlm.nih.gov/30700729/>), where cogsci students (in dataset 1) and schizophrenia patients + controls (dataset 2) combine their own intuition of trustworthiness of given faces to social information.

Your task is to

implement a simulation of the task
implement 2 different agents playing the task in R and the relative Stan inferential models (WBA, the other is of your choice, e.g. SBA, PBA, or another of your liking).
design 3-5 scenarios(i. e. different weights - one relying on social information, one not caring, some in between) that distinguish the agent and visualize the behaviors 
simulate data from the models to assess whether the models can be fit and the true model can be inferred via model comparison
    has the sampling worked? divergences, etc
    model recovery on the 3-5 scenarios
    [predictive checks; prior posterior updates; prior sensitivity, SBC and parameter recovery are optional]
    [multilevel is optional]
fit the 2 models to one empirical dataset (if you choose to leave multilevel out, you are fitting a no pooling model, each participant on its own]
    data exploration
    divergences, etc
    predictive checks
    prior posterior updates
    [prior/likelihood sensitivity and multilevel are optional]
    model comparison
report (v minimal description of research question, v minimal description of data, description of models, model quality checks, report of results, both of model comparison and of how the parameter values look like)

```{r}
# Packages
pacman::p_load(
  tidyverse,    # Data manipulation and plotting
  here,         # Robust file paths
  cmdstanr,     # Stan interface
  posterior,    # Posterior analysis
  bayesplot,    # Plotting MCMC
  furrr,        # Parallel processing
  future,        # Future backend
  loo,
  patchwork
)

# parallel setup
n_cores <- parallel::detectCores() -2
plan(multisession, workers = n_cores)

#default ggplot theme
theme_set(theme_classic())
```


```{r}
exp_data <- read_csv("assignment3_data.csv") # data supplied by Riccardo
```

```{r}
# Create the full factorial design
evidence_combinations <- expand_grid(
  rating_1  = 0:7,   # Direct evidence: 0-8 blue marbles out of 8
  social  = 0:7,   # Social evidence: encoded as 0-3 (pseudo-marbles)
 # total1 = 8,     # Fixed sample size for direct evidence
 # total2 = 3      # Fixed "sample size" for social evidence
)

# This gives us 9 × 4 = 36 unique evidence combinations
cat("Unique evidence combinations:", nrow(evidence_combinations), "\n")
```

```{r}
# Core belief-update function implementing the Beta-Binomial integration model
# for trustworthiness ratings on a 0-7 scale
betaBinomialTrust <- function(alpha_prior = 0.5, beta_prior = 0.5,
                               rating_1, social,
                               w_direct = 1, w_social = 1) {
  
  alpha_post    <- alpha_prior + w_direct * rating_1 + w_social * social
  beta_post     <- beta_prior  + w_direct * (7 - rating_1) + w_social * (7 - social)
  expected_rate <- alpha_post / (alpha_post + beta_post)
  ci_lower      <- qbeta(0.025, alpha_post, beta_post)
  ci_upper      <- qbeta(0.975, alpha_post, beta_post)
  
  list(
    alpha_post    = alpha_post,
    beta_post     = beta_post,
    expected_rate = expected_rate,
    ci_lower      = ci_lower,
    ci_upper      = ci_upper,
    # final rating: rescale expected_rate back to 1-8 scale
    choice        = round(expected_rate * 7)
  )
}

# Function to generate decisions from a simulated agent
generate_agent_decisions_trust <- function(weight_direct, weight_social,
                                            evidence_df, n_samples = 5) {
  evidence_df |>
    slice(rep(seq_len(n()), each = n_samples)) |>
    group_by(rating_1, social) |>
    mutate(sample_id = seq_len(n())) |>
    ungroup() |>
    pmap_dfr(function(rating_1, social, sample_id) {
      result <- betaBinomialTrust(
        0.5, 0.5,
        rating_1  = rating_1,
        social    = social,
        w_direct  = weight_direct,
        w_social  = weight_social
      )
      # Probability matching: sample theta from Beta, then sample choice from Binomial(7, theta)
      alpha_p <- result$alpha_post
      beta_p  <- result$beta_post
      p       <- rbeta(1, alpha_p, beta_p)
      choice  <- rbinom(1, 7, p)   # draw on 0-7 scale
      
      tibble(
        sample_id     = sample_id,
        rating_1      = rating_1,
        social        = social,
        expected_rate = result$expected_rate,
        choice        = choice,                  # 0-7 scale for Stan
        choice_label  = choice + 1               # 1-8 scale for reporting
      )
    })
}

# Prepare Stan-friendly data list
prepare_stan_data_trust <- function(df) {
  list(
    N             = nrow(df),
    choice        = df$choice,        # 0-7 scale
    rating_1      = df$rating_1,      # 0-7 scale
    social        = df$social,        # 0-7 scale
    prior_rho_alpha   = 2.0,
    prior_rho_beta    = 2.0,
    prior_kappa_mu    = 0.5,
    prior_kappa_sigma = 0.5,
    run_diagnostics   = 1
  )
}
```


```{r}
# Scenario: agent rates trustworthiness 5/8, group rates 7/8
# On 0-7 scale: rating_1 = 4, social = 6
rating_1_demo <- 4
social_demo   <- 6

alpha_demo <- 0.5 + rating_1_demo + social_demo        # = 10.5
beta_demo  <- 0.5 + (7 - rating_1_demo) + (7 - social_demo)  # = 4.5
n_sims     <- 10000

# Each simulated agent makes n_ratings trustworthiness judgments
n_ratings  <- 20
expected_rate_demo <- alpha_demo / (alpha_demo + beta_demo)

# ── Step 1: Simulate both models ───────────────────────────────────────────────

# Model A: Beta-Binomial (uncertainty in theta propagates)
# Draw theta from Beta, then draw ratings based on that theta
theta_draws <- rbeta(n_sims, alpha_demo, beta_demo)
bb_draws    <- rbinom(n_sims, size = 7, prob = theta_draws)  # single rating draw on 0-7

# Model B: Standard Binomial (fixed belief, no uncertainty in theta)
binom_draws <- rbinom(n_sims, size = 7, prob = expected_rate_demo)

# ── Step 2: Visualization ──────────────────────────────────────────────────────

# Plot 1: uncertainty in theta
p_theta <- tibble(theta = theta_draws) |>
  ggplot(aes(x = theta)) +
  geom_density(fill = "#4575b4", alpha = 0.5, color = "transparent") +
  geom_vline(xintercept = expected_rate_demo,
             colour = "#d73027", linewidth = 1, linetype = "dashed") +
  scale_x_continuous(limits = c(0, 1), labels = scales::percent) +
  labs(
    title    = "Step 1: Uncertainty in trustworthiness belief (θ)",
    subtitle = sprintf("Beta(%.1f, %.1f) | Dashed line is E[θ] = %.0f%%\nrating_1 = %d, social = %d (0-7 scale)",
                       alpha_demo, beta_demo, expected_rate_demo * 100,
                       rating_1_demo, social_demo),
    x = expression("Latent trustworthiness tendency (" * theta * ")"),
    y = "Density"
  )

# Plot 2: outcome comparison showing overdispersion
# Shift draws to 1-8 scale for interpretability
plot_data <- tibble(
  `Beta-Binomial (Uncertain θ)` = bb_draws + 1,
  `Standard Binomial (Fixed θ)` = binom_draws + 1
) |>
  pivot_longer(cols = everything(), names_to = "Model", values_to = "Rating")

p_spread <- ggplot(plot_data, aes(x = Rating, fill = Model)) +
  geom_bar(position = "identity", alpha = 0.6, aes(y = after_stat(prop))) +
  scale_fill_manual(values = c("Standard Binomial (Fixed θ)" = "#d73027",
                               "Beta-Binomial (Uncertain θ)" = "#4575b4")) +
  scale_x_continuous(breaks = 1:8) +
  scale_y_continuous(labels = scales::percent) +
  labs(
    title    = "Step 2: Trustworthiness rating distribution",
    subtitle = "Beta-Binomial shows wider spread — uncertainty in θ propagates to ratings",
    x        = "Trustworthiness rating (1-8)",
    y        = "Probability",
    fill     = NULL
  ) +
  theme(legend.position = "top")

# ── Combine panels ─────────────────────────────────────────────────────────────
p_theta + p_spread +
  plot_annotation(
    title    = "How uncertainty propagates: Beta-Binomial vs Binomial",
    subtitle = sprintf(
      "Agent's own rating: %d | group's rating: %d (0-7 scale)\nWhen belief about trustworthiness is uncertain, ratings spread more widely.",
      rating_1_demo, social_demo)
  )
```

```{r}
# All combinations of agent rating and group rating on 0-7 scale
evidence_combinations_trust <- expand.grid(
  rating_1 = 0:7,
  social   = 0:7
)

# Calculate SBA predictions for all evidence combinations
sim_data_trust <- evidence_combinations_trust |>
  pmap_dfr(function(rating_1, social) {
    result <- betaBinomialTrust(0.5, 0.5,
                                rating_1 = rating_1,
                                social   = social)
    tibble(
      rating_1      = rating_1,
      social        = social,
      expected_rate = result$expected_rate,
      ci_lower      = result$ci_lower,
      ci_upper      = result$ci_upper
    )
  }) |>
  mutate(
    social_evidence = factor(
      social, levels = 0:7,
      labels = c("1 - Very low",  "2 - Low",
                 "3 - Below mid", "4 - Above mid",
                 "5 - Moderate",  "6 - High",
                 "7 - Very high", "8 - Maximum")
    )
  )

semantic_colors <- c(
  "1 - Very low"  = "#d73027",
  "2 - Low"       = "#f46d43",
  "3 - Below mid" = "#fdae61",
  "4 - Above mid" = "#fee090",
  "5 - Moderate"  = "#abd9e9",
  "6 - High"      = "#74add1",
  "7 - Very high" = "#4575b4",
  "8 - Maximum"   = "#313695"
)

ggplot(sim_data_trust,
       aes(x = rating_1 + 1, y = expected_rate,
           colour = social_evidence, group = social_evidence)) +
  geom_ribbon(aes(ymin = ci_lower, ymax = ci_upper,
                  fill = social_evidence), alpha = 0.15, colour = NA) +
  geom_line(linewidth = 1) +
  geom_point(size = 3) +
  geom_hline(yintercept = 0.5, linetype = "dashed", colour = "grey50") +
  scale_x_continuous(breaks = 1:8) +
  scale_colour_manual(values = semantic_colors) +
  scale_fill_manual(values   = semantic_colors) +
  coord_cartesian(ylim = c(0, 1)) +
  labs(
    title    = "Simple Bayesian Agent: trustworthiness evidence integration",
    subtitle = "Expected trustworthiness tendency with 95% credible intervals | Weights fixed at (1, 1)",
    x        = "Agent's own initial rating (1-8 scale)",
    y        = "Expected trustworthiness tendency (θ)",
    colour   = "group rating",
    fill     = "group rating"
  ) +
  theme(legend.position = "bottom")
```

```{r}
N_predict <- 10

overdispersion_data_trust <- expand_grid(
  rating_1 = 0:7,
  social   = 0:7
) |>
  mutate(
    alpha_p   = 0.5 + rating_1 + social,
    beta_p    = 0.5 + (7 - rating_1) + (7 - social),
    theta_hat = alpha_p / (alpha_p + beta_p),

    # Expected Binomial variance for N draws (baseline)
    var_binom = N_predict * theta_hat * (1 - theta_hat),

    # Beta-Binomial variance for N draws (includes belief uncertainty)
    var_bb    = var_binom * (1 + (N_predict - 1) / (alpha_p + beta_p + 1)),

    # Absolute excess variance: Var(BetaBinomial) - Var(Binomial)
    excess_variance = var_bb - var_binom,

    social_evidence = factor(
      social, levels = 0:7,
      labels = c("1 - Very low",  "2 - Low",
                 "3 - Below mid", "4 - Above mid",
                 "5 - Moderate",  "6 - High",
                 "7 - Very high", "8 - Maximum")
    )
  )

ggplot(overdispersion_data_trust,
       aes(x = rating_1 + 1, y = excess_variance,
           colour = social_evidence, group = social_evidence)) +

  geom_hline(yintercept = 0, linetype = "dashed", color = "gray70") +
  geom_line(linewidth = 1) +
  geom_point(size = 2) +
  scale_colour_manual(values = semantic_colors) +
  scale_x_continuous(breaks = 1:8) +

  labs(
    title    = "Beta-Binomial overdispersion across trustworthiness evidence combinations",
    subtitle = paste(
      "Excess variance = Var(BetaBinomial) − Var(Binomial) for N=10 predicted draws.",
      "\nHighest when own rating and group rating pull in opposite directions."
    ),
    x        = "Agent's own initial rating (1-8 scale)",
    y        = "Excess variance\n(driven by belief uncertainty)",
    colour   = "group rating"
  ) +
  theme(
    legend.position = "bottom",
    plot.title      = element_text(face = "bold", size = 14),
    axis.title.y    = element_text(face = "bold", margin = margin(r = 10)),
    axis.title.x    = element_text(face = "bold", margin = margin(t = 10))
  )
```

```{r}
agents <- tibble(
  agent_type    = c("Self-Focused", "Balanced", "Socially-Influenced"),
  weight_direct = c(1.5, 1.0, 0.5),
  weight_social = c(0.5, 1.0, 2.0)
)

# Generate simulated data for all three agent types
simulation_results_trust <- agents |>
  pmap_dfr(function(agent_type, weight_direct, weight_social) {
    generate_agent_decisions_trust(weight_direct, weight_social,
                                   evidence_combinations_trust,
                                   n_samples = 5) |>
      mutate(agent_type = agent_type)
  }) |>
  mutate(
    social_evidence = factor(
      social, levels = 0:7,
      labels = c("1 - Very low",  "2 - Low",
                 "3 - Below mid", "4 - Above mid",
                 "5 - Moderate",  "6 - High",
                 "7 - Very high", "8 - Maximum")
    ),
    agent_type = factor(
      agent_type,
      levels = c("Self-Focused", "Balanced", "Socially-Influenced")
    )
  )

# Summarize for plotting
plot_summary_trust <- simulation_results_trust |>
  group_by(agent_type, rating_1, social_evidence) |>
  summarise(
    theoretical_p = mean(expected_rate),
    empirical_p   = mean(choice / 7),   # rescale 0-7 to 0-1 for comparability
    .groups = "drop"
  )

# Visualize
ggplot(plot_summary_trust,
       aes(x = rating_1 + 1, colour = social_evidence, group = social_evidence)) +
  geom_line(aes(y = theoretical_p), linewidth = 1.2, alpha = 0.5) +
  geom_line(aes(y = empirical_p), linetype = "dashed", linewidth = 0.8) +
  geom_point(aes(y = empirical_p), size = 2) +
  geom_hline(yintercept = 0.5, linetype = "dotted", colour = "grey30") +
  facet_wrap(~ agent_type) +
  scale_colour_manual(values = semantic_colors) +
  scale_x_continuous(breaks = 1:8) +
  coord_cartesian(ylim = c(0, 1)) +
  labs(
    title    = "Weighted Bayesian agents: three conformity scenarios",
    subtitle = "Solid faded = internal belief | Dashed = observed rating proportions (N=5 per condition)",
    x        = "Agent's own initial rating (1-8 scale)",
    y        = "Expected trustworthiness tendency (θ)",
    colour   = "group rating"
  ) +
  theme(legend.position  = "bottom",
        strip.background = element_blank(),
        strip.text       = element_text(face = "bold", size = 11))
```

```{r}
SimpleAgent_stan <- "
// Simple Bayesian Agent (SBA) — Face Trustworthiness Rating.
// No free parameters — evidence is counted at face value.
// Jeffreys prior pseudo-counts: alpha0 = beta0 = 0.5.
// Ratings are on a 0-7 scale (original 1-8 shifted by -1 in R).
data {
  int<lower=1> N;
  array[N] int<lower=0, upper=7> choice;  // agent's final rating (0-7)
  array[N] int<lower=0, upper=7> rating_1; // agent's own initial rating (0-7)
  array[N] int<lower=0, upper=7> social;   // group's rating (0-7)
  int<lower=0, upper=1> run_diagnostics;
}
transformed data {
  real alpha0 = 0.5;
  real beta0  = 0.5;
}
model {
  // Vectorized likelihood — fixed weights = 1
  vector[N] alpha_post = alpha0 + to_vector(rating_1) + to_vector(social);
  vector[N] beta_post  = beta0
                         + (7 - to_vector(rating_1))
                         + (7 - to_vector(social));
  target += beta_binomial_lpmf(choice | 7, alpha_post, beta_post);
}
generated quantities {
  vector[N] log_lik;
  array[N] int prior_pred;
  array[N] int posterior_pred;

  if (run_diagnostics) {
    for (n in 1:N) {
      real alpha_post = alpha0 + rating_1[n] + social[n];
      real beta_post  = beta0  + (7 - rating_1[n]) + (7 - social[n]);

      log_lik[n]        = beta_binomial_lpmf(choice[n] | 7, alpha_post, beta_post);
      prior_pred[n]     = beta_binomial_rng(7, alpha0, beta0);
      posterior_pred[n] = beta_binomial_rng(7, alpha_post, beta_post);
    }
  }
}
"
write_stan_file(SimpleAgent_stan,
                dir      = "stan/",
                basename = "sba_trust.stan")
```

```{r}
WeightedAgent_stan <- "
// Weighted Bayesian Agent (reparameterised) — Face Trustworthiness Rating.
// rho in (0,1): relative weight of direct vs social evidence.
// kappa > 0: total evidence scaling (w_d + w_s).
// Ratings are on a 0-7 scale (original 1-8 shifted by -1 in R).
data {
  int<lower=1> N;
  array[N] int<lower=0, upper=7> choice;   // agent's final rating (0-7)
  array[N] int<lower=0, upper=7> rating_1; // agent's own initial rating (0-7)
  array[N] int<lower=0, upper=7> social;   // group's rating (0-7)

  real prior_rho_alpha;
  real prior_rho_beta;
  real prior_kappa_mu;
  real<lower=0> prior_kappa_sigma;
  int<lower=0, upper=1> run_diagnostics;
}
parameters {
  real<lower=0, upper=1> rho;   // relative weight: w_d / (w_d + w_s)
  real<lower=0>          kappa; // total weight: w_d + w_s
}
transformed parameters {
  real<lower=0> weight_direct = rho * kappa;
  real<lower=0> weight_social = (1.0 - rho) * kappa;
}
model {
  // rho: weakly centred on equal weighting
  target += beta_lpdf(rho | prior_rho_alpha, prior_rho_beta);
  // kappa: lognormal centered on 2 (SBA equivalent when rho = 0.5)
  target += lognormal_lpdf(kappa | prior_kappa_mu, prior_kappa_sigma);

  // Vectorized likelihood
  vector[N] alpha_post = 0.5 + weight_direct * to_vector(rating_1)
                             + weight_social * to_vector(social);
  vector[N] beta_post  = 0.5 + weight_direct * (7 - to_vector(rating_1))
                             + weight_social * (7 - to_vector(social));
  target += beta_binomial_lpmf(choice | 7, alpha_post, beta_post);
}
generated quantities {
  vector[N] log_lik;
  array[N] int prior_pred;
  array[N] int posterior_pred;
  real lprior = beta_lpdf(rho | prior_rho_alpha, prior_rho_beta) +
                lognormal_lpdf(kappa | prior_kappa_mu, prior_kappa_sigma);

  // Draw from priors for prior predictive
  real rho_prior   = beta_rng(prior_rho_alpha, prior_rho_beta);
  real kappa_prior = lognormal_rng(prior_kappa_mu, prior_kappa_sigma);
  real wd_prior    = rho_prior * kappa_prior;
  real ws_prior    = (1.0 - rho_prior) * kappa_prior;

  if (run_diagnostics) {
    for (n in 1:N) {
      // Posterior quantities
      real alpha_post = 0.5
        + weight_direct * rating_1[n]
        + weight_social * social[n];
      real beta_post  = 0.5
        + weight_direct * (7 - rating_1[n])
        + weight_social * (7 - social[n]);

      // Prior quantities
      real alpha_prior = 0.5
        + wd_prior * rating_1[n]
        + ws_prior * social[n];
      real beta_prior  = 0.5
        + wd_prior * (7 - rating_1[n])
        + ws_prior * (7 - social[n]);

      log_lik[n]        = beta_binomial_lpmf(choice[n] | 7, alpha_post, beta_post);
      prior_pred[n]     = beta_binomial_rng(7, alpha_prior, beta_prior);
      posterior_pred[n] = beta_binomial_rng(7, alpha_post,  beta_post);
    }
  }
}
"
write_stan_file(WeightedAgent_stan,
                dir      = "stan/",
                basename = "wba_trust.stan")
```


```{r}
sba_model <- cmdstan_model("stan/sba_trust.stan")
wba_model <- cmdstan_model("stan/wba_trust.stan")

sampler_args <- list(
  chains = 4, 
  parallel_chains = 4, 
  iter_warmup = 1000, 
  iter_sampling = 1000, 
  refresh = 0
)
```

```{r}
# True values for the three WBA scenarios
# wd and ws map to weight_direct and weight_social
# which in turn determine rho = wd/(wd+ws) and kappa = wd+ws
true_wba <- list(
  Self_Focused        = list(wd = 1.5, ws = 0.5),  # rho = 0.75, kappa = 2
  Balanced            = list(wd = 1.0, ws = 1.0),  # rho = 0.50, kappa = 2
  Socially_Influenced = list(wd = 0.5, ws = 2.0)   # rho = 0.20, kappa = 2.5
)

# Generate simulated data for each scenario
stan_data_wba <- map(true_wba, function(params) {
  sim <- generate_agent_decisions_trust(
    params$wd, params$ws,
    evidence_combinations_trust,
    n_samples = 5
  )
  prepare_stan_data_trust(sim)
})
```

```{r}
dir.create(here::here("simmodels"), showWarnings = FALSE)
```


```{r}
fit_sba_paths <- here::here(
  "simmodels",
  paste0("sba_trust_", names(stan_data_wba), ".rds")
) |> set_names(names(stan_data_wba))

regenerate_simulations <- TRUE

if (regenerate_simulations || !all(file.exists(fit_sba_paths))) {
  fit_sba <- imap(stan_data_wba, function(data, scenario) {
    fit <- sba_model$sample(
      data          = data,
      chains        = 2,
      iter_warmup   = 1000,
      iter_sampling = 2000,
      fixed_param   = FALSE,
      refresh       = 500,
      seed          = 123
    )
    fit$save_object(fit_sba_paths[[scenario]])
    fit
  })
} else {
  fit_sba <- map(fit_sba_paths, readRDS)
}
```

```{r}
fit_wba_paths <- here::here(
  "simmodels",
  paste0("wba_trust_", names(stan_data_wba), ".rds")
) |> set_names(names(stan_data_wba))

if (regenerate_simulations || !all(file.exists(fit_wba_paths))) {
  seeds <- set_names(101:103, names(stan_data_wba))

  fit_wba <- imap(stan_data_wba, function(data, scenario) {
    fit <- wba_model$sample(
      data            = data,
      seed            = seeds[[scenario]],
      chains          = 4,
      parallel_chains = 4,
      iter_warmup     = 1000,
      iter_sampling   = 2000,
      refresh         = 500
    )
    fit$save_object(fit_wba_paths[[scenario]])
    fit
  })
} else {
  fit_wba <- map(fit_wba_paths, readRDS)
}
```


```{r}
# Generalized extraction function for posteriors
extract_predictive_data_trust <- function(fit, stan_data, scenario_name,
                                          target_var = "posterior_pred") {
  obs_df <- tibble(
    obs_id   = seq_len(stan_data$N),
    rating_1 = stan_data$rating_1,
    social   = stan_data$social,
    choice   = stan_data$choice
  )

  pred_samples <- posterior::as_draws_df(fit$draws(target_var)) |>
    dplyr::select(-.chain, -.iteration) |>
    tidyr::pivot_longer(
      cols      = dplyr::starts_with(target_var),
      names_to  = "obs_id",
      values_to = "pred"
    ) |>
    dplyr::mutate(obs_id = readr::parse_number(obs_id))

  obs_summary <- obs_df |>
    dplyr::group_by(rating_1, social) |>
    dplyr::summarise(
      p_obs = mean(choice / 7),   # rescale 0-7 to 0-1 for comparability
      .groups = "drop"
    )

  pred_summary <- pred_samples |>
    dplyr::left_join(obs_df, by = "obs_id") |>
    dplyr::group_by(.draw, rating_1, social) |>
    dplyr::summarise(p_draw = mean(pred / 7), .groups = "drop") |>
    dplyr::group_by(rating_1, social) |>
    dplyr::summarise(
      p_pred = median(p_draw),
      lower  = quantile(p_draw, 0.025),
      upper  = quantile(p_draw, 0.975),
      .groups = "drop"
    ) |>
    dplyr::left_join(obs_summary, by = c("rating_1", "social")) |>
    dplyr::mutate(
      social_evidence = factor(
        social, levels = 0:7,
        labels = c("1 - Very low",  "2 - Low",
                   "3 - Below mid", "4 - Above mid",
                   "5 - Moderate",  "6 - High",
                   "7 - Very high", "8 - Maximum")
      ),
      Scenario = scenario_name
    )

  return(pred_summary)
}

# Standardized plotting function
plot_predictive_check_trust <- function(plot_data, title, subtitle) {

  ggplot(plot_data,
         aes(x = rating_1 + 1, colour = social_evidence,
             fill = social_evidence)) +
    geom_linerange(aes(ymin = lower, ymax = upper),
                   position = position_dodge(width = 0.6),
                   alpha = 0.5, linewidth = 2) +
    geom_line(aes(y = p_pred),
              position = position_dodge(width = 0.6),
              linewidth = 1) +
    geom_point(aes(y = p_obs),
               shape = 21, colour = "black", stroke = 0.8, size = 2,
               position = position_dodge(width = 0.6)) +
    facet_wrap(~ Scenario) +
    scale_color_manual(values = semantic_colors) +
    scale_fill_manual(values  = semantic_colors) +
    scale_x_continuous(breaks = 1:8) +
    coord_cartesian(ylim = c(0, 1)) +
    labs(
      title    = title,
      subtitle = subtitle,
      x        = "Agent's own initial rating (1-8 scale)",
      y        = "Expected trustworthiness tendency (θ)",
      colour   = "group rating",
      fill     = "group rating"
    ) +
    theme_bw() +
    theme(legend.position = "bottom",
          strip.text = element_text(face = "bold", size = 11))
}

# Extract SBA posterior predictive checks across all three scenarios
sba_post_ppc <- imap_dfr(fit_sba, function(fit, scenario) {
  extract_predictive_data_trust(
    fit           = fit,
    stan_data     = stan_data_wba[[scenario]],
    scenario_name = scenario,
    target_var    = "posterior_pred"
  )
})

# Plot SBA across all three scenarios
plot_predictive_check_trust(
  sba_post_ppc,
  title    = "Posterior predictive checks: Simple Bayesian Agent",
  subtitle = "SBA fits the Balanced scenario but misses Self-Focused and Socially-Influenced systematically"
)
```

```{r}
n_sims <- 50

dir.create(here::here("simmodels/temp"), showWarnings = FALSE, recursive = TRUE)

if (regenerate_simulations || !file.exists(here::here("simmodels/model_recovery_trust.rds"))) {

  # Generate datasets from each model type
  sim_grid <- bind_rows(
    tibble(sim_id = 1:n_sims, true_model = "SBA", wd = 1.0, ws = 1.0),
    tibble(sim_id = 1:n_sims, true_model = "WBA",
           rho   = rbeta(n_sims, 2, 2),
           kappa = rlnorm(n_sims, log(2), 0.5)) |>
      mutate(wd = rho * kappa, ws = (1 - rho) * kappa)
  )

  recovery_data <- sim_grid |>
    mutate(
      agent_data = pmap(list(wd, ws), function(wd, ws) {
        generate_agent_decisions_trust(wd, ws,
                                       evidence_combinations_trust,
                                       n_samples = 5)
      })
    )

  results_list <- list()

  for (i in seq_len(nrow(recovery_data))) {
    id       <- recovery_data$sim_id[[i]]
    true_mod <- recovery_data$true_model[[i]]
    df       <- recovery_data$agent_data[[i]]
    true_wd  <- recovery_data$wd[[i]]
    true_ws  <- recovery_data$ws[[i]]

    stan_data <- prepare_stan_data_trust(df)

    # Fit both models
    # Fit SBA with fixed_param
fit_s <- sba_model$sample(
  data          = stan_data,
  chains        = 2,
  iter_warmup   = 0,
  iter_sampling = 500,
  fixed_param   = TRUE,
  refresh       = 0,
  show_messages = FALSE,
  output_dir    = here::here("simmodels/temp")
)

fit_w <- wba_model$sample(
  data          = stan_data,
  chains        = 2,
  iter_warmup   = 1000,
  iter_sampling = 1000,
  refresh       = 0,
  show_messages = FALSE,
  output_dir    = here::here("simmodels/temp")
)

# SBA: compute ELPD directly from log_lik (no LOO needed - no parameters)
sba_log_lik <- fit_s$draws("log_lik", format = "matrix")
sba_elpd    <- sum(colMeans(sba_log_lik))

# WBA: use LOO normally
loo_w    <- fit_w$loo()
wba_elpd <- loo_w$estimates["elpd_loo", "Estimate"]

# Compare via ELPD difference instead of stacking weights
elpd_diff  <- wba_elpd - sba_elpd
winner     <- ifelse(elpd_diff > 0, "WBA", "SBA")

wba_draws <- fit_w$draws(format = "df")

results_list[[i]] <- tibble(
  sim_id     = id,
  true_model = true_mod,
  true_wd    = true_wd,
  true_ws    = true_ws,
  sba_elpd   = sba_elpd,
  wba_elpd   = wba_elpd,
  elpd_diff  = elpd_diff,   # positive = WBA wins
  winner     = winner,
  wba_wd     = mean(wba_draws$rho * wba_draws$kappa),
  wba_ws     = mean((1 - wba_draws$rho) * wba_draws$kappa)
)
  }

  recovery_results <- bind_rows(results_list)
  saveRDS(recovery_results, here::here("simmodels/model_recovery_trust.rds"))

} else {
  recovery_results <- readRDS(here::here("simmodels/model_recovery_trust.rds"))
}

# ── Visualize model recovery ───────────────────────────────────────────────────
recovery_results |>
  ggplot(aes(x = true_model, y = elpd_diff, fill = true_model)) +
  geom_boxplot(alpha = 0.7) +
  geom_hline(yintercept = 0, linetype = "dashed", colour = "grey40") +
  scale_fill_manual(values = c("SBA" = "#5B8DB8", "WBA" = "#D4735A")) +
  labs(
    title    = "Model recovery: ELPD difference (WBA - SBA) by true generating model",
    subtitle = "Positive = WBA preferred | Negative = SBA preferred\nCorrect recovery: SBA data → negative, WBA data → positive",
    x        = "True generating model",
    y        = "ELPD difference (WBA - SBA)",
    fill     = "True model"
  ) +
  theme(legend.position = "none",
        strip.text = element_text(face = "bold", size = 11))
```



