---
title: "ACM Assignment 3"
author: "Sara Szabo"
date: "2026-04-23"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Task description

In assignment 3 you have to analyze real world data using Bayesian models of cognition. You can apply the models we discussed during the lectures, but you need to adjust them to reflect the specific setup.

The data comes from the social conformity experiment (<https://pubmed.ncbi.nlm.nih.gov/30700729/>), where cogsci students (in dataset 1) and schizophrenia patients + controls (dataset 2) combine their own intuition of trustworthiness of given faces to social information.

Your task is to

implement a simulation of the task
implement 2 different agents playing the task in R and the relative Stan inferential models (WBA, the other is of your choice, e.g. SBA, PBA, or another of your liking).
design 3-5 scenarios(i. e. different weights - one relying on social information, one not caring, some in between) that distinguish the agent and visualize the behaviors 
simulate data from the models to assess whether the models can be fit and the true model can be inferred via model comparison
    has the sampling worked? divergences, etc
    model recovery on the 3-5 scenarios
    [predictive checks; prior posterior updates; prior sensitivity, SBC and parameter recovery are optional]
    [multilevel is optional]
fit the 2 models to one empirical dataset (if you choose to leave multilevel out, you are fitting a no pooling model, each participant on its own]
    data exploration
    divergences, etc
    predictive checks
    prior posterior updates
    [prior/likelihood sensitivity and multilevel are optional]
    model comparison
report (v minimal description of research question, v minimal description of data, description of models, model quality checks, report of results, both of model comparison and of how the parameter values look like)

```{r}
# Packages
pacman::p_load(
  tidyverse,    # Data manipulation and plotting
  here,         # Robust file paths
  cmdstanr,     # Stan interface
  posterior,    # Posterior analysis
  bayesplot,    # Plotting MCMC
  furrr,        # Parallel processing
  future        # Future backend
)

# parallel setup
n_cores <- parallel::detectCores() -2
plan(multisession, workers = n_cores)

#default ggplot theme
theme_set(theme_classic())
```


```{r}
exp_data <- read_csv("assignment3_data.csv") # data supplied by Riccardo
```
# Simulating task


```{r}
set.seed(123)

n_agents <- 100 
n_trials <- 10 
rating_scale <- 8 # generate quantity between 1 and 8
alpha_0 <- rep(0.5, rating_scale) # Jeffreys prior
 
true_values <- sample(1:rating_scale, n_trials, replace = TRUE) # only here for simulation's sake
group_values <- sample(1:rating_scale, n_trials, replace = TRUE) # second source of information

# sampler function - draws 1 rating given a Dirichlet posterior

draw_rating <- function(alpha_post){ 
  theta <- rgamma(length(alpha_post), shape = alpha_post, rate = 1)
  theta <- theta / sum(theta)
  sample(seq_along(alpha_post), size = 1, prob = theta)
}

one_agent_one_trial <- function(id, true_value, group_value, alpha_0, rating_scale, w = 1) {
  
  # drawing the agent's own initial rating
  init_counts <- rep(0, rating_scale)
  init_counts[true_value] <- 1
  
  # agent meets social input
  social_counts <- rep(0, rating_scale)
  social_counts[group_value] <- 1
  
  rating_1 <- draw_rating(alpha_0 + init_counts)
  alpha_post <- alpha_0 + init_counts + w * social_counts # weight is fixed at 1
  rating_2 <- draw_rating(alpha_post)
  
  tibble(
    agent_id = id,
    rating_1 = rating_1,
    social_input = group_value,      
    rating_2 = rating_2,            
    weight = w
  )
}

# multiple agents & multiple trials

simulate_SBA <- function(n_agents, n_trials, true_values, group_values, alpha_0, rating_scale, w =1){ # multiple agents & multiple trials
  map_dfr(1:n_trials, function(trial){
  map_dfr(1:n_agents, one_agent_one_trial,
          true_value = true_values[trial],
          group_value = group_values[trial],
          alpha_0 = alpha_0,
          rating_scale = rating_scale,
          w = 1) %>% 
      mutate(trial = trial)
  })
}

sba_data <- simulate_SBA(
  n_agents = n_agents,
  n_trials = n_trials, 
  true_values = true_values,
  group_values = group_values,
  alpha_0 = alpha_0,
  rating_scale = rating_scale, 
  w = 1
)

stan_data <- list(
  N_total = nrow(sba_data),
  rating_scale = rating_scale,
  rating_1 = sba_data$rating_1,
  group_rating = sba_data$social_input,
  y = sba_data$rating_2
)


# Introducing WBA

# defining multiple scenarios
scenarios <- list(
  "low conformity"     = 0.25,
  "moderate conformity"           = 0.50,
  "equal to SBA"       = 1.00,
  "high conformity"    = 2.00,
  "very high conformity"= 4.00
)

# generates a single trial output for both agents
both_agents_single_trial <- function(id, true_value, group_value, alpha_0, rating_scale, w) {

  init_counts               <- rep(0, rating_scale)
  init_counts[true_value]  <- 1

  social_counts                     <- rep(0, rating_scale)
  social_counts[group_value] <- 1

  rating_1 <- draw_rating(alpha_0 + init_counts)

  # SBA - w = 1 fixed
  alpha_sba  <- alpha_0 + init_counts + 1 * social_counts
  final_sba  <- draw_rating(alpha_sba)

  # WBA - w from scenarios
  alpha_wba  <- alpha_0 + init_counts + w * social_counts
  final_wba  <- draw_rating(alpha_wba)

  tibble( # for different values stored, make changes here
    agent_id = id,
    rating_1 = rating_1,
    social_input = group_value,
    final_sba  = final_sba,
    change_sba = final_sba - rating_1,
    final_wba = final_wba,
    change_wba = final_wba - rating_1,
    weight       = w
  )
}



# different WBA scenarios + SBA data
both_agents_data <- imap_dfr(scenarios, function(w, label) {
  map_dfr(1:n_trials, function(trial){
    map_dfr(1:n_agents, both_agents_single_trial,
            true_value = true_values[trial],
            group_value = group_values[trial],
            alpha_0 = alpha_0,
            rating_scale = rating_scale,
            w = w) %>% 
      mutate(trial = trial)
  }) %>% 
    mutate(scenario = label)
})

# generating stan-friendly format (1 list per scenario)
both_agents_stan_data <- imap(scenarios, function(w, label) { # for different values stored, make changes here
  results <- both_agents_data %>% filter(scenario == label)
  list(
    N_total = nrow(results),
    rating_scale = rating_scale,
    rating_1 = results$rating_1,
    group_rating = results$social_input,
    y_sba = results$final_sba, # final rating by SBA
    y_wba = results$final_wba, # final raring by WBA
    change_sba = results$change_sba,
    change_wba = results$change_wba
  )
})

# summary
both_agents_data %>%
  group_by(scenario) %>%
  summarise(
    mean_initial = mean(rating_1),
    mean_sba = mean(final_sba),
    mean_wba = mean(final_wba),
    change_sba = mean(final_sba - rating_1),
    change_wba = mean(final_wba - rating_1)
  ) %>%
  print()


```


# Agents & Stan models

# Simulate data from models

# Fitting models to empirical dataset