---
title: "Assignment 1"
author: "Sara Szabo"
date: "2026-02-12"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
# install.packages(c( "pacman", "remotes", "tidyverse", "data.table", "dtplyr", "downlit", "xml2", "rmarkdown", "sessioninfo", "brms" ))
```

# Based on notebook

```{r}
# Load necessary packages
# patchwork helps combine multiple plots
pacman::p_load(tidyverse, patchwork)

# Number of trials per simulation
trials <- 120  

# Number of agents to simulate
agents <- 100  

# Optional: Set random seed for reproducibility
set.seed(123)
```

```{r}
# --- Simulation Parameters ---
trials <- 120 # Number of trials for a single simulation run

# --- Basic Implementation ---
# Simulate an unbiased random agent (rate = 0.5)
rate_unbiased <- 0.5
choices_unbiased <- rbinom(trials, size = 1, prob = rate_unbiased)

# rbinom generates random draws from a binomial distribution.
# size = 1 makes it a Bernoulli trial (0 or 1 outcome).
# prob = rate specifies the probability of getting a 1.

# Simulate a biased random agent (rate = 0.8, chooses 'right' 80% of the time)
rate_biased <- 0.8
choices_biased <- rbinom(trials, size = 1, prob = rate_biased)

# --- Visualization 1: Raw Choices ---
# Create data frames for plotting
d_unbiased <- tibble(trial = 1:trials, choice = choices_unbiased, rate = rate_unbiased)
d_biased <- tibble(trial = 1:trials, choice = choices_biased, rate = rate_biased)

# Plot raw choice sequence (0s and 1s)
p1 <- ggplot(d_unbiased, aes(x = trial, y = choice)) +
  geom_line(color = "blue") + geom_point(color = "blue", size = 1) +
  labs(title = "Unbiased Random Agent (Rate = 0.5)", y = "Choice (0/1)") + theme_classic() + ylim(-0.1, 1.1)

p2 <- ggplot(d_biased, aes(x = trial, y = choice)) +
  geom_line(color = "red") + geom_point(color = "red", size = 1) +
  labs(title = "Biased Random Agent (Rate = 0.8)", y = "Choice (0/1)") + theme_classic() + ylim(-0.1, 1.1)

# Show plots side-by-side
print(p1 + p2 + plot_layout(guides = "collect"))
```

# Assignment description

[insert assignment description here]

# Based on slides

```{r}
# Unbiased Random Agent

RandomAgent_f <- function(rate){
  choice <- rbinom(1, 1, rate) # specifying mechanism + decision making
  return(choice) # output
}
```


```{r}
rate <- 0.5 # rate (theta) - choocing 1 this % of times
randomChoice <- RandomAgent_f(rate)
randomChoice
```


```{r}
trials <- 120 # how many trials?

randomChoice <- rep(NA, trials) # creating empty vector
for (t in seq(trials)){
  randomChoice[t] <- RandomAgent_f(rate) # running fucntion as many times as the trials & saving output
}

d1 <- tibble(trial = seq(trials), choice = randomChoice) # making a df
p1 <- ggplot(d1, aes(trial, choice)) + 
  geom_line() +
  theme_classic() # plot

p1 # display the plot
```


```{r}
d1 <- NULL # making sure the df is not there already

for (rate in seq(0, 1, 0.1)){# looping through rates of 0, 0.1, 0.2..., up to 1
  randomChoice <- rep(NA, trials)
  for (t in seq(trials)){randomChoice[t] <- RandomAgent_f(rate)}
  temp <- tibble(trial = seq(trials), choice = randomChoice, rate)
  if(exists("d1")){d1 <- rbind(d1, temp)} 
  else{d1 <- temp}
}

p1 <- ggplot(d1, aes(trial, choice)) + geom_line()+facet_wrap(.~rate)+
  theme_classic()
p1

d1 <- d1 %>% 
  group_by(rate) %>% 
  mutate(
    cumulative_rate = cumsum(choice)/seq_along(choice)
  )
```


```{r}
p3 <- ggplot(d1, aes(trial, cumulative_rate))+
  geom_line()+
  geom_hline(yintercept = 0.5, linetype = "dashed")+
  ylim(0,1)+
  facet_wrap(.~rate)+
  theme_classic()

p3
```


```{r}
p4 <- ggplot(d1, aes(trial, cumulative_rate,
               color = rate, group = rate))+
  geom_line()+
  geom_hline(yintercept = 0.5,
             linetype = "dashed")+
  ylim(0,1)+
  theme_classic()

p4
```
### Random Agent with noise
```{r}
RandomAgentNoise_f <- function(rate, noise){
  choice <- rbinom(1, 1, rate)
  if (rbinom(1, 1, noise)==1){choice = rbinom(1, 1, 0.5)}
  return(choice)
}
```

Implementing noise
```{r}
d2 <- NULL

for (noise in seq(0, 0.5, 0.1)){
  for (rate in seq(0, 1, 0.1)){
    randomChoice <- rep(NA, trials)
    for (t in seq(trials)){randomChoice[t] <- RandomAgentNoise_f(rate, noise)}
    temp <- tibble(trial = seq(trials), choice = randomChoice, rate, noise)
    if (exists("d2")){d2 <- rbind(d2, temp)} 
    else{d2 <- temp}
  }
}

d2 <- d2 %>% 
  group_by(rate, noise) %>% 
  mutate(cumulative_rate = cumsum(choice)/seq_along(choice)
         )

p4 <- ggplot(d2, aes(trial, cumulative_rate, group = rate, color = rate))+
  geom_line()+
  geom_hline(yintercept = 0.5, linetype = "dashed")+
  ylim(0,1)+
  facet_wrap(.~noise)+
  theme_classic()

p4

```

# Win-stay-lose-shift Agent

```{r}
WSLSAgent_f <- function(prevChoice, feedback){
  if (feedback == 1){
    choice = prevChoice}
  else if (feedback == 0){
    choice = 1-prevChoice}
  return(choice)
}
```

```{r}
WSLSChoice <- WSLSAgent_f(1, 0)
```

```{r}
# against random agent
rate <- 0.7

Self <- rep(NA, trials)
Other <- rep(NA, trials)

Self[1] <- RandomAgent_f(0.5)

for (t in seq(trials)){Other[t] <- RandomAgent_f(rate)}

for (i in 2:trials){
  if (Self[i-1]== Other[i-1]){ # Self = Matcher
    feedback = 1
  }
  else {feedback == 0}
  Self[i] <- WSLSAgent_f(Self[i-1], feedback)
}
```


```{r}
# visualizing it
df <- tibble(Self, Other, trial = seq(trials), Feedback = as.numeric(Self==Other))

ggplot(df)+theme_classic()+geom_line(color = "red", aes(trial, Self)) +
  geom_line(color = "blue", aes(trial, Other))
```

```{r}
df <- df %>% mutate(
  cumulative_rateSelf = cumsum(feedback) / seq_along(feedback),
  cumulative_rateOther = cumsum(1-feedback) / seq_along(feedback)
)

ggplot(df) + theme_classic() + geom_line(color = "red", aes(trial, cumulative_rateSelf)) + geom_line(color = "blue", aes(trial, cumulative_rateOther))
```

## Note: get back to slide 36 (wsls vs wsls)

Idea: RL (switching based on wins/losses) vs WSLS
biased agent is also an option

Question: is it probabilistic, stochastic, deterministic?

Our agent will fight against WSLS and random agent

Idea: change strategy midway -> e deterministic until 50% of trials, then what?


# Implementing reinforcement learning

$$  V_t = V_{t-1} + \alpha \cdot (Feedback_t - V_{t-1}) = V_{t-1} + \alpha \cdot PE$$

Current value = weighted sum of all past rewards where weight decays exponentially

$$Weight_k = \alpha(1-\alpha)^k  $$ 

```{r}
# for example

WSLSAgent_f <- function(prevChoice, feedback){
  if (feedback == 1){
    choice = prevChoice}
  else if (feedback == 0){
    choice = 1-prevChoice}
  return(choice)
}

WSLSChoice <- WSLSAgent_f(1, 0)


WSLSAgent_f <- function(prevChoice, feedback, noise = 0) {
  # Input validation
  if (!prevChoice %in% c(0, 1)) stop("Previous choice must be 0 or 1.")
  if (!feedback %in% c(0, 1)) stop("Feedback must be 0 or 1.")
  if (!is.numeric(noise) || noise < 0 || noise > 1) stop("Noise must be a probability between 0 and 1.")

  # Core WSLS logic:
  # If feedback is 1 (win), stay: choice = prevChoice
  # If feedback is 0 (loss), shift: choice = 1 - prevChoice
  choice <- ifelse(feedback == 1, prevChoice, 1 - prevChoice)

  # Apply noise if specified
  if (noise > 0 && runif(1) < noise) {
    # Override with a random 50/50 choice
    choice <- sample(c(0, 1), 1)
  }

  return(choice)
}
```



```{r}
# previous choice = previous expected value
# reward = feedback

trials = 120
V = rep(NA, trials)

RL_agent_f <- function(count, alpha, reward){
  #PE = 
  V[count] <<- V[count-1] + alpha * (reward - V[count-1])
  choice = rbinom(1,1,V[count])
  #choice <- ifelse(expected_value >= 0.5, 1, 0) # this is probably wrong, I need to understand how it works
  return(choice) # has to be binary - the one piece I don't have yet
}
```

```{r}
RL_agent_f(trials, 0.8, 0)
```


# Simulating choices of this thing
```{r}
Self = rep(NA, trials)
Other = rep(NA, trials)
result = rep(NA, trials)
V[1] = 0.5
biased_agent_rate = 0.7

alpha = 0.5

Self[1] = RL_agent_f(trials, alpha, result)

for(t in seq(trials)){
  Other[t] = RandomAgent_f(biased_agent_rate)
}
Other
Self

```


```{r}
alpha <- 0.3
noise <- 0.05

# initializing
Self[1] <- RandomAgentNoise_f(biased_agent_rate, noise)

result[1] <- NA

#result[1] = if(Other[1]==Self[1]){feedback = 1}else{feedback = 0}

for(i in 2:trials){
  if(Self[i-1] == Other[i-1]){
    reward = Self[i-1]
  }
  else{
    reward = Other[i-1]
  }
  Self[i] = RL_agent_f(i, alpha, reward)
  Other[i] = RandomAgentNoise_f(biased_agent_rate, noise)
  result[i] = reward
}
Self
result
```



```{r}
df = tibble(Self, Other, trial = seq(trials), Feedback = as.numeric(Self==Other))

ggplot(df) + 
  theme_classic() + 
  geom_line(color = "red", aes(trial,Self), size = 1) + 
  geom_line(color = "blue", aes(trial,Other), size = 1) + 
  geom_line(color="green", aes(trial,V), size = 1.5) + 
  geom_hline(yintercept = biased_agent_rate, size = 1.5)
```

```{r}
d9 <- d9 %>% 
  group_by(rate) %>% 
  mutate(
    cumulative_rate = cumsum(choice)/seq_along(choice)
  )

p9 <- ggplot(d9, aes(trial, cumulative_rate,
               color = rate, group = rate))+
  geom_line()+
  geom_hline(yintercept = 0.5,
             linetype = "dashed")+
  ylim(0,1)+
  theme_classic()

p4
```







Note: does the model act according to the assumptions I put into it? Compare performance against existing implementation

# add exponential decay of memory