In the assignment you need to build and validate a cognitive model of matching pennies behavior using Stan. As an outcome of this you will have to produce a text (+ plots) document (linked to a github repo) in which you:
  
  describe the model you are working on (N.B. it doesn't need to be one from Assignment 1)
   - showcase a commented version of the stan model (what does each line do?)
   - discuss model quality of the model (predictive prior and posterior checks, prior-posterior update checks, etc.)
   - describe a process of parameter recovery (why are you doing it?, how are you doing it?)
   - discuss the results: can you recover the parameter values? How many trials should be used at least to properly recover the parameters? What’s the role of priors? Add relevant plot(s).

N.B. to pass it's enough to use the single agent model, but if you want to learn more (at the cost of more time investment), you could
                                         
(optional) Build the multilevel version of your model and run the parameter recovery on it.
(optional) Analyze an empirical dataset
                                         
                                         Empirical datasets:
                                           
Guide to the data: https://www.dropbox.com/s/m1u5roo4z0f2emr/readme.txt?dl=0
Non-human primates dataset: https://www.dropbox.com/s/8fj7k5jk8mn4j63/mp_primates.csv?dl=0
Human primates dataset (former CogSci'ers):
 https://www.dropbox.com/s/9z6hcg9b4yhzgbu/mp_students.csv?dl=0
 https://www.dropbox.com/s/r2ok23vjewyelaz/mp_students_22.csv?dl=0
Clinical dataset (schizophrenia and controls): https://www.dropbox.com/s/dxj9f8txg1yax68/mp_schizophrenia.csv?dl=0

## Packages 
```{r}
pacman::p_load(
  tidyverse,    # Data manipulation and plotting
  here,         # Robust file paths
  cmdstanr,     # Stan interface
  posterior,    # Posterior analysis
  bayesplot,    # Plotting MCMC
  furrr,        # Parallel processing
  future        # Future backend
)

# parallel setup
n_cores <- parallel::detectCores() -2
plan(multisession, workers = n_cores)

#default ggplot theme
theme_set(theme_classic())

```

# Simulating data

```{r simulating data}
# Trials and agents from Yosuf's code

set.seed(123)
trials <-  120

WSLSAgent_f=function(prevChoice, feedback){
  if(feedback == 1){
    choice = prevChoice
  }
  else if (feedback == 0) {
    choice = 1 -prevChoice 
  }
return(choice) 
}

# true values for later recovery
set.seed(123)
trials <-  120
true_alpha <- 0.3
true_beta  <- 2.0

# initialize all vectors
Other    <- rep(NA, trials)
Self     <- rep(NA, trials)
feedback <- rep(NA, trials)
V        <- rep(NA, trials)

# initializing with everything starting at 0.5
Other[1]    <- rbinom(1, 1, 0.5)
V[1]        <- 0.5
Self[1]     <- rbinom(1, 1, V[1])
feedback[1] <- as.integer(Self[1] == Other[1])


for (t in 2:trials) {
  Other[t] <- WSLSAgent_f(Other[t-1], feedback[t-1]) # WSLS agent starts
  V[t]     <- V[t-1] + true_alpha * (Other[t-1] - V[t-1]) 
  p        <- 1 / (1 + exp(-true_beta * (V[t] - 0.5)))
  Self[t]  <- rbinom(1, 1, p) # probability of choosing 1
  
  feedback[t] <- as.integer(Self[t] == Other[t])
}


```

```{r}

```


```{r}
# # simulate one agent with known params
# # "true" values for later recovery
# 
# true_alpha <- 0.3
# true_beta <- 2.0
# 
# # opponent - WSLS agent
# Other    <- rep(NA, trials)
# Other[1] <- rbinom(1, 1, 0.5)
# for (t in 2:trials) {
#   Other[t] <- WSLSAgent_f(Other[t-1], feedback[t-1])
# }
# 
# # Self - RL agent
# Self <-  rep(NA, trials)
# # Other  <- rep(NA, trials)
# # memory <- rep(NA,trials)
# # result <- rep(NA,trials)
# V <- rep(NA,trials)
# V[1] <- 0.5
# Self[1] <- rbinom(1, 1, V[1])
# 
# for (t in 2:trials){
#   V[t] <- V[t-1] + true_alpha * (Other[t-1] - V[t-1])
#   # softmax
#   p <- 1/(1 + exp(-true_beta * (V[t] - 0.5)))
#   Self[t] <- rbinom(1, 1, p)
# }
# 
# # feedback
# feedback <- as.integer(Self == Other)
```

```{r data-prep}
# preparing data for STAN
data_for_stan <- list(
  n = trials,
  players = 1L, 
  choice = Self, # agent's choices - 0 or 1
  feedback = feedback,
  prior_alpha_m = 0,
  prior_alpha_sd = 1,
  prior_beta_m = 0, 
  prior_beta_sd = 1
)
```

```{r}
data_for_stan <- list(
  n = trials,
  Self  = Self,
  Other = Other
)
```




# STAN model

*for now this is all from the slides*
```{r}
stan_file_RL <- here("Assignment_2",  "model_outline.stan")

mod_RL <- cmdstan_model(
  stan_file_RL,
  cpp_options = list(stan_threads = FALSE),
  stanc_options = list("O1")
  )
```

```{r}
stan_file_RL <- here("Assignment_2",  "RL.stan")

mod_RL <- cmdstan_model(
  stan_file_RL,
  cpp_options = list(stan_threads = FALSE),
  stanc_options = list("O1")
  )
```





```{r fitting-model}
# Sample from the posterior distribution using MCMC
fit_RL <- mod_RL$sample(
  data = data_for_stan, # should make this
  seed = 123, # reproducibility
  chains = 4, # nr. of MArkov-chains
  parallel_chains = 4, #run them in parallel
  # threads_per_chain = 1, # within-chain parallelization, here just in case
  iter_warmup = 1000, # get discarded
  iter_sampling = 2000, 
  refresh = 500, # how often should progress updates be printed
  max_treedepth = 10, # complexity of mcmc steps
  adapt_delta = 0.99 # target acceptance rate
  ) 
```


# Model quality checks

```{r}
# summary of params
fit_RL$summary()

# divergences
fit_RL$diagnostic_summary()

mcmc_trace(fit_RL$draws(), pars = c("alpha", "beta"))
```

```{r extract-draws}
# converting posterior draws to tidy df
draws_df <- fit_RL$draws(format = "df")

#glimpse(draws_df)

```


## Generating prior and posterior predictions
- priors
- simulations from the prior
- simulations from the posterior

To do: 
- separate for alpha and beta once the structure is more clarified
- find a way to feed more educated guesses to the model


## Prior-posterior update checks



```{r}
# # RICCARDO
# 
# alpha
ggplot(draws_df)+
  geom_density(aes(alpha), fill = "blue", alpha = 0.3) + 
  geom_density(aes(alpha_prior), fill = "red", alpha = 0.3) +
  geom_vline(xintercept = true_alpha, color='red')+
  labs(
    title = "Prior-Posterior update for alpha",
    subtitle = "red = prior, blue = posterior, vertical line = True alpha",
    x = "alpha", y = "Density"
  ) +
  theme_classic()

# beta 
ggplot(draws_df)+
  geom_density(aes(beta), fill = "blue", alpha = 0.3) + 
  geom_density(aes(beta_prior), fill = "red", alpha = 0.3) +
  geom_vline(xintercept = true_beta, color='red')+
  scale_x_continuous(limits = c(0,10))+
  labs(
    title = "Prior-Posterior update for beta",
    subtitle = "red = prior, blue = posterior, vertical line = True beta",
    x = "beta", y = "Density"
  ) +
  theme_classic()
```

## Prior predictive check

```{r}

prior_pred_sums <- draws_df %>% 
  select(starts_with("choice_prior_pred")) %>% 
  rowSums()

# RICCARDO

# prior predictive check
ggplot() +
  geom_histogram(aes(prior_pred_sums), color = "darkblue", fill = "blue",
                 alpha = 0.3, bins = 90) +
  geom_point(x = sum(Self), y = 0, color = "red", shape = 17, size = 5) +
  xlab("Predicted right hands out of 120 trials")+
  ylab("Count")+
  theme_classic()

```


## Posterior predictive checks


```{r}
posterior_pred_sums <- draws_df %>%
  select(starts_with("Self_rep")) %>%
  rowSums()

ggplot() +
  geom_histogram(aes(posterior_pred_sums),
                 color = "darkblue",
                 fill = "blue",
                 alpha = 0.3,
                 bins = 90) +
  labs(
    title = "Posterior predictive check",
    x = "Predicted right hands out of 120 trials", y = "Count"
  ) +
  theme_classic()



```




```{r}

# sum across trials for each draw — gives "how many heads predicted per draw"
prior_pred_sums <- draws_df %>% 
  select(starts_with("choice_prior_pred")) %>% 
  rowSums()

posterior_pred_sums <- draws_df %>% 
  select(starts_with("choice_posterior_pred")) %>% 
  rowSums()

# RICCARDO

# alpha

ggplot(draws_df)+
  geom_histogram(aes(prior_pred_sums),
                 color = "lightblue", fill = "blue",
                 alpha = 0.3, bins = 90) +
  geom_histogram(aes(posterior_pred_sums),
                 color = "darkblue", fill = "blue",
                 alpha = 0.3, bins = 90)+
  geom_point(x = sum(Self), y = 0, color = "red",
             shape = 17, size = 5) +
  xlab("Predicted heads out of 120 trials") +
  ylab("Density")+
  theme_classic()

```


# Parameter recovery

```{r}
# recovery fun for a single alpha-beta pair

recover_params <- function(true_alpha, true_beta, trials = 120, seed = NULL) {
  if (!is.null(seed)) set.seed(seed)
  
  # initialising vectors
  V <- rep(NA, trials) # p of choosing 1
  choice <- rep(NA, trials)
  Other <- rep(NA, trials)
  feedback <- rep(NA, trials)
  
  # first trial
  Other[1] <- rbinom(1, 1, 0.5)
  V[1] <- 0.5
  choice[1] <- rbinom(1, 1, V[1])
  feedback[1] <- as.integer(choice[1]==Other[1]) # should there be weights for this? This is just T/F

# rest of trials
  for (t in 2:trials) {
    Other[t]       <- WSLSAgent_f(Other[t-1], feedback[t-1])
    V[t]           <- V[t-1] + true_alpha * (Other[t-1] - V[t-1])
    p              <- 1 / (1 + exp(-true_beta * (V[t] - 0.5)))
    choice[t]      <- rbinom(1, 1, p)
    feedback[t]<- as.integer(choice[t] == Other[t])
  }
  
  feedback <- as.integer(choice[t] == Other[t])
  
  # Fit model
  stan_data <- list(
    n = trials,
    players = 1L, # 1 as integer
    choice = choice,
    feedback = feedback,
    prior_alpha_m = 0, prior_alpha_sd = 1, # both normal distributions
    prior_beta_m = 0, prior_beta_sd = 1
  )
  
  fit <- mod_RL$sample(
    data = stan_data,
    seed = 123,
    chains = 4,
    parallel_chains = 4,
    iter_warmup = 1000,
    iter_sampling = 2000,
    refresh = 0, # silent for recovery loop
    adapt_delta = 0.99
  )
  
  # get posterior means and 90% credible intervals
  summ <- fit$summary(c("alpha", "beta"))
  
  tibble(
    true_alpha = true_alpha,
    true_beta = true_beta,
    est_alpha = summ$mean[summ$variable == "alpha"],
    est_beta = summ$mean[summ$variable == "beta"],
    lo_alpha = summ$q5[summ$variable == "alpha"],
    hi_alpha = summ$q95[summ$variable == "alpha"],
    lo_beta = summ$q5[summ$variable == "beta"],
    hi_beta = summ$q95[summ$variable == "beta"],
    
  )
}

# grid of param combinations to recover
recovery_grid <- expand_grid(
  true_alpha = c(0.1, 0.3, 0.5, 0.7, 0.9),
  true_beta = c(0.5, 1.0, 2.0, 4.0)
)

# running recovery in parallel
recovery_results <- #future_
pmap_dfr(
  recovery_grid,
  ~ recover_params(..1, ..2, trials = 500),
  .options = furrr_options(seed = TRUE)
)

```

```{r}
recover_params <- function(true_alpha, true_beta, trials = 120, seed = NULL) {
  if (!is.null(seed)) set.seed(seed)
  
  # storage
  V <- rep(NA, trials) # p of choosing 1
  choice <- rep(NA, trials)
  Other <- rep(NA, trials)
  
  # first trial
  Other[1] <- rbinom(1, 1, 0.5)
  V[1] <- 0.5
  p1 <- plogis(true_beta * (V[1] - 0.5))
  choice[1] <- rbinom(1, 1, p1)
  
  # remaining trials
  for (t in 2:trials) {
    Other[t] <- WSLSAgent_f(Other[t-1], choice[t-1] == Other[t-1])
    V[t] <- V[t-1] + true_alpha * (Other[t-1] - V[t-1])
    p <- plogis(true_beta * (V[t] - 0.5))
    choice[t] <- rbinom(1, 1, p)
  }
  
  # Stan data
  stan_data <- list(
    n = trials,
    Self = choice,
    Other = Other
  )
  
  # fit model
  fit <- mod_RL$sample(
    data = stan_data,
    seed = 123,
    chains = 4,
    parallel_chains = 4,
    iter_warmup = 1000,
    iter_sampling = 2000,
    refresh = 0,
    adapt_delta = 0.99
  )
  
  # extract posterior means and intervals
  summ <- fit$summary(c("alpha", "beta"))
  
  tibble(
    trials = trials,
    true_alpha = true_alpha,
    true_beta = true_beta,
    est_alpha = summ$mean[summ$variable == "alpha"],
    est_beta = summ$mean[summ$variable == "beta"],
    lo_alpha = summ$q5[summ$variable == "alpha"],
    hi_alpha = summ$q95[summ$variable == "alpha"],
    lo_beta = summ$q5[summ$variable == "beta"],
    hi_beta = summ$q95[summ$variable == "beta"]
  )
}
```

```{r}

recovery_grid <- expand_grid(
  true_alpha = c(0.1, 0.3, 0.5, 0.7, 0.9),
  true_beta = c(0.5, 1.0, 2.0, 4.0)
)

recovery_results <- pmap_dfr(
  recovery_grid,
  ~ recover_params(..1, ..2, trials = 500),
  .options = furrr_options(seed = TRUE)
)


```





```{r}
# plotting param recov results

# alpha 
ggplot(recovery_results, aes(true_alpha, est_alpha)) +
  geom_pointrange(aes(ymin = lo_alpha, ymax = hi_alpha), alpha = 0.6) +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "red") +
  labs(
    title = "Parameter recovery: alpha",
    subtitle = "Points = posterior mean, bars = 90% CI, dashed = perfect recovery",
    x = "True alpha", y = "Estimated alpha"
  )

# beta
ggplot(recovery_results, aes(true_beta, est_beta)) +
  geom_pointrange(aes(ymin = lo_beta, ymax = hi_beta), alpha = 0.6) +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "red") +
  labs(
    title = "Parameter recovery: beta",
    subtitle = "Points = posterior mean, bars = 90% CI, dashed = perfect recovery",
    x = "True beta", y = "Estimated beta"
  )
```


```{r}

# Recovery with different amount of trials 
trial_lengths <- c(50, 100, 200, 400, 800)

recovery_grid <- expand_grid(
  trials = trial_lengths,
  true_alpha = c(0.1, 0.3, 0.5, 0.7, 0.9),
  true_beta = c(0.5, 1, 2, 4)
)

recovery_results <- pmap_dfr(
  recovery_grid,
  ~ recover_params(..2, ..3, trials = ..1)
)



```


```{r}
recovery_results %>%
  group_by(trials) %>%
  summarise(
    rmse_alpha = sqrt(mean((est_alpha - true_alpha)^2)),
    rmse_beta  = sqrt(mean((est_beta - true_beta)^2))
  ) %>%
  pivot_longer(cols = starts_with("rmse")) %>%
  ggplot(aes(x = trials, y = value, color = name)) +
  geom_line(size = 1.2) +
  geom_point(size = 3) +
  ylab("RMSE") + xlab("Number of trials") +
  theme_classic()
```



```{r}

ggplot(recovery_results, aes(true_alpha, est_alpha)) +
  geom_pointrange(aes(ymin = lo_alpha, ymax = hi_alpha), alpha = 0.6) +
  geom_abline(slope = 1, intercept = 0, 
              linetype = "dashed", color = "red") +
  facet_wrap(~ trials, nrow = 2) +
  labs(
    title = "Parameter recovery: α (learning rate)",
    subtitle = "Each panel shows a different number of trials\nPoints = posterior mean, bars = 90% CI, dashed = perfect recovery",
    x = "True α",
    y = "Estimated α"
  ) +
  theme_classic()

ggplot(recovery_results, aes(true_beta, est_beta)) +
  geom_pointrange(aes(ymin = lo_beta, ymax = hi_beta), alpha = 0.6) +
  geom_abline(slope = 1, intercept = 0, 
              linetype = "dashed", color = "red") +
  facet_wrap(~ trials, nrow = 2) +
  labs(
    title = "Parameter recovery: β (inverse temperature)",
    subtitle = "Each panel shows a different number of trials\nPoints = posterior mean, bars = 90% CI, dashed = perfect recovery",
    x = "True β",
    y = "Estimated β"
  ) +
  theme_classic()

```



